---
title: "PA1_template"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Load data and convert the date field
```{r CACHE=TRUE}
#read data from the activity.csv file and load it into activity 
        activity<-read.csv("activity.csv",header=TRUE,dec=".")
#convert date field to date format
        activity$date<-as.Date(activity$date)  
```
#Different plottings  
##Histogram of the total number of steps each day
```{r  CACHE=TRUE}
#histogram about sum of steps taken by day
        pro_activity<-aggregate(steps~date,activity,sum)
        hist(pro_activity$steps,xlab="Sum of steps taken per day",ylab="number of days") 
```  

##Histogram of the mean number of steps each day   
```{r CACHE=TRUE}
 #histogram about mean of steps taken by day
        pro_activity<-aggregate(steps~date,activity,mean)
        hist(pro_activity$steps,xlab="Mean of steps taken per day",ylab="number of days")  
```  

##Histogram of the median number of steps each day
```{r CACHE=TRUE}
#histogram about median of steps taken by day
       pro_activity<-aggregate(activity,by=list(activity$steps,activity$date),FUN=median,na.rm=TRUE)
        hist(pro_activity$steps,xlab="Median of steps taken per day",ylab="number of days")  
```  

##Time series plot of the average number of steps taken
```{r CACHE=TRUE}
#time series based average step take per day
        pro_activity<-aggregate(steps~date,activity,mean)
        with(pro_activity,plot(date,steps))  
```  

##The interval that contains on average basis the maximum number of steps
```{r CACHE=TRUE}
#The 5 min interval that, on average, contains the maximum of steps
        interval_activity<-aggregate(steps~interval,activity,mean)
        x<-interval_activity[interval_activity$steps==max(interval_activity$steps),1]  
```
###The 5 min-interval containing the maximum average value is `r x`  

##Imputting missing values  
After identying the rows containing steps with missing values. Creation of a new table with those missing values will be replaced by the average value of the corresponding interval.  

*Select NA values and store them in a data.table called test*

```{r CACHE=TRUE}
#select NA values
test<- activity[is.na(activity$steps),]
```  

*adapt in the test subset the interval so no more NA fields*
```{r CACHE=TRUE}
#adapt in the test subset the interval so no more NA fields
 test[,1]<-interval_activity[test$interval==interval_activity$interval,2]
```  

*cloning activity to new_activity and paste the result in the new activity data frame*
```{r CACHE=TRUE}
#paste the result in the new activity data frame
new_activity<-activity
new_activity[is.na(new_activity$steps),1]<-test[new_activity$date==test$date & new_activity$interval==test$interval,1]
``` 

##Histogram of the total number of steps each day after missing values imputted
```{r  CACHE=TRUE}
#histogram about sum of steps taken by day
        pro_activity<-aggregate(steps~date,new_activity,sum)
        hist(pro_activity$steps,xlab="Sum of steps taken per day",ylab="number of days") 
```  

##Histogram of the mean number of steps each day after missing values imputted
```{r  CACHE=TRUE}
#histogram about sum of steps taken by day
        pro_activity<-aggregate(steps~date,new_activity,mean)
        hist(pro_activity$steps,xlab="Mean of steps taken per day",ylab="number of days") 
```  

##Histogram of the median number of steps each day after missing values imputted
```{r  CACHE=TRUE}
#histogram about sum of steps taken by day
       pro_activity<-aggregate(new_activity,by=list(new_activity$steps,new_activity$date),FUN=median,na.rm=TRUE)
       hist(pro_activity$steps,xlab="Median of steps taken per day",ylab="number of days")
``` 

*Filling the NA values made figures higher*  


*To compare weekdays and weekend activity let's create 2 temporary table frame one for weekdays and the other for weekends*

```{r  CACHE=TRUE}
#load the dplyr library because needed now
library(dplyr)
#creation of weekends table
we_activity<-new_activity%>%filter(weekdays(date)=="samedi"|weekdays(date)=="dimanche")%>%mutate(weekend="Weekend")

#creation of weekdays table
wd_activity<-new_activity%>%filter(weekdays(date)!="samedi"& weekdays(date)!="dimanche")%>%mutate(weekend="Weekday")

#creation of aggregate tables for weekends and weekdays
wd_activity<-aggregate(steps~interval,wd_activity,mean)
we_activity<-aggregate(steps~interval,we_activity,mean)

#merge data into total_activity
we_activity<-we_activity%>%mutate(days="weekend")
wd_activity<-wd_activity%>%mutate(days="weekday")
total_activity<-rbind(wd_activity,we_activity)

#plot result
library(ggplot2)
ggplot(data=total_activity,aes(x=interval,y=steps,colour=days))+ggtitle("Comparison between average activity during the week days and weekend")+geom_line()
```  


